\documentclass[11pt, oneside]{article} 
\usepackage{geometry}
\geometry{letterpaper} 
\usepackage{graphicx}
	
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{parskip}
\usepackage{color}
\usepackage{hyperref}

\graphicspath{{/Users/telliott/Github/figures/}}

\title{Laurent series cheatsheet}
\date{}

\begin{document}
\maketitle
\Large

%[my-super-duper-separator]

\label{sec:series_cheatsheet}

In general, the Laurent series is two parts.  The analytic part is
\[ \sum a_k (z - z_0)^k \]
valid in a disk inside $r = 1$.

The principal part is
\[ \sum b_k (z - z_0)^{-k} \]
valid in the region outside a disk of radius $r = 1$.  If the variable is $a/z$ then the valid radius is $z > a$.

Dealing with Laurent series problems is a pain.  I always seem to be re-doing the same calculations, hence, this cheatsheet.

\subsection*{before you start}

First 

$\circ$ \ Change of variables to move the contour to the origin.

$\circ$ \ Partial fractions decomposition, if possible.

$\circ$ \ Change of sign overall to fit the forms we develop below.

$\circ$ \ For $1/(1 + z)$ substitute $-z$ for $z$.

\subsection*{basic series}

$A$:
\[ \frac{1}{1 - z} = 1 + z + z^2 + z^3 \dots \]
converges for an open disk with $|z| < 1$.

Use the trick for getting a \emph{different} geometric series.  Write
\[ = -\frac{1}{z} \cdot \frac{1}{1 - 1/z} \]

That's also a geometric series, but in $1/z$.  It converges when $|1/z| < 1$.

But then $|z| > 1$!  That's just what we need to work \emph{outward} from a radius.  $P$ (for principal) is $-1/z$ times the geometric series in $1/z$:

$P$:
\[ \frac{1}{1 - z} = -\frac{1}{z} \cdot (1 + \frac{1}{z} + \frac{1}{z^2} + \frac{1}{z^3} \dots ) \]
\[ \frac{1}{1 - z} = - \frac{1}{z} - \frac{1}{z^2} - \frac{1}{z^3}  - \frac{1}{z^4} \dots \]

We can read the value of the integral as the coefficient $b_1 = -1$.

\subsection*{modification}

There are three other simple forms with a constant other than $1$:

\subsection*{$1/a + z$}

\label{sec:a_plus_z}

\[ \frac{1}{a + z} = \frac{1}{a} \cdot \frac{1}{1 + z/a} \]
Write the $A$ series in $-z/a$:
\[ \frac{1}{a} \ [ \ 1 -\frac{z}{a} + (\frac{z}{a})^2 - (\frac{z}{a})^3 \dots \ ] \]
For the $P$ series, factor out $1/z$
\[ \frac{1}{a + z} = \frac{1}{z} \cdot \frac{1}{1 + a/z} \]
and write the $P$ series in $-a/z$:
\[ \frac{1}{z} \cdot \ [ \ 1 - \frac{a}{z} + (\frac{a}{z})^2 - (\frac{a}{z})^3 \dots \ ] \]
\[ \frac{1}{z} - \frac{a}{z^2} + (\frac{a}{z})^2 - (\frac{a}{z})^3 \dots  \]

\subsection*{$1/a - z$}

\label{sec:a_minus_z}

For the second one we have the analytic series in terms of $z/a$ 
\[ \frac{1}{a - z} = \frac{1}{a} \cdot \frac{1}{1 - z/a} = \frac{1}{a} \cdot \ [ \ 1 +  \frac{z}{a} + (\frac{z}{a})^2 + \dots \ ] \]
For the $P$ series we need a different factorization:
\[ \frac{1}{a - z} = -\frac{1}{z} \cdot \frac{1}{1 - a/z} \]
\[ = -\frac{1}{z} \ [ 1 +  \frac{a}{z} + (\frac{a}{z})^2 + (\frac{a}{z})^3 \dots \ ]  \]
\[ = -\frac{1}{z} - \frac{a}{z^2} - \frac{a^2}{z^3} - \frac{a^3}{z^4} \dots  \]

The third is minus the second.

\label{sec:z_minus_a}

\subsection*{corollaries}

\[ \frac{1}{1 + z^2} = 1 + (-z^2) + (-z^2)^2 + (-z^2)^3 \dots \]

If you have something like
\[ \frac{z}{1 - z} \]

that's just $z$ times the appropriate series above.

\end{document}