\documentclass[11pt, oneside]{article} 
\usepackage{geometry}
\geometry{letterpaper} 
\usepackage{graphicx}
	
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{parskip}
\usepackage{color}
\usepackage{hyperref}

\graphicspath{{/Users/telliott/Github/figures/}}

\title{Polar CRE}
\date{}

\begin{document}
\maketitle
\Large

In this chapter we will derive the CRE conditions for polar coordinates and show how to compute the derivative in the same system.  The take home lesson is that there is an extra factor for each:

Recall that the CRE in Cartesian coordinates are:
\[ u_x = v_y \]
\[ v_x = -u_y \]
It turns out that $r,\theta$ is similar except for a factor of $r$, which goes on the partial with respect to $r$
\[ r u_r = v_{\theta} \]
\[ r v_r = - u_{\theta} \]

Also,
\[ f'(z) = e^{-i \theta} \ (u_r + i v_r) \]

We'll also see two derivations for each, a simple one and a more careful but also more complicated one. 

A clever way to derive the CRE in polar coordinates is to take advantage of the result that we obtained in Cartesian coordinates.

\subsection*{short and sweet}

The function $f(z) = z = x + iy$ is "analytic" and obeys the CRE.  

Write the same function in polar coordinates:
\[ z = re^{i \theta} \]
and then separate it into a completely real part $u(r, \theta)$ and a completely imaginary part $v(r,\theta)$:
\[ z = r(\cos \theta + i \sin \theta) \]
\[ = r \cos \theta + i r \sin \theta \]
Observe that
\[ u_r = \cos \theta \]
\[ v_{\theta} = r \cos \theta \]
so we deduce that
\[ r u_r =  v_{\theta} \]

Similarly
\[ v_r = \sin \theta \]
\[ u_{\theta} = - r \sin \theta \]
so
\[ -r v_r = u_{\theta} \]
There are the CRE in polar coordinates.  Carrying out the same computation for \emph{any} analytic function would give the same result (with some other expression in the middle, of course).

\[ r u_r = v_{\theta} \]
\[ r v_r = - u_{\theta} \]

Notice the similar format to the Cartesian version, with the addition of a factor of $r$.  

It reminds me of the Jacobian from multi-variable calculus.

\subsection*{polar derivative} 
This will also turn out to be very similar to the Cartesian version, with an extra factor out front:
\[ f'(z) = e^{-i \theta} \ [ \ u_r + i v_r \ ] \]

Let's just assume that the derivative is equal to what we would expect, within some unknown factor of $k$:
\[ z' = k(u_r + i v_r) \]
and now we know that for this function the derivative is equal to $1$:
\[ f(z) = z \]
\[ z' = 1 = k(u_r + i v_r) \]
if we write in polar coordinates:
\[ z = r \cos \theta + i r \sin \theta \]
Then 
\[ u_r + i v_r = \cos \theta + i \sin \theta \]
What is the factor that multiplies this expression to give $1$?
Clearly
\[ e^{- i \theta} (\cos \theta + i \sin \theta) = 1 \]
So $k = e^{- i \theta}$.

\subsection*{CRE derivation by the chain rule}

We know equations to go back and forth between $x,y$ and $r, \theta$ so it is not hard to imagine that we can always re-write $u$ and $v$ as
\[ z = u \ [ \ x(r, \theta), y(r, \theta) \ ] + i v \ [ \ x(r,\theta),y(r, \theta) \ ] \]
or more succinctly:
\[ z = u(r, \theta) + i v(r, \theta) \]

Now we ask about relations between the partial derivatives.  Let us first make a table of them:
\[ x = r \cos \theta \]
\[ x_r = \cos \theta, \ \ \ \ \ \ x_{\theta} = -r \sin \theta \]
\[ y = r \sin \theta \]
\[ y_r = \sin \theta, \ \ \ \ \ \ y_{\theta} = r \cos \theta \]

Clearly, we want expressions involving $u_r$, $v_{\theta}$ etc.  Write:
\[ \frac{\partial u}{\partial r} = \frac{\partial u}{\partial x} \frac{\partial x}{\partial r} + \frac{\partial u}{\partial y} \frac{\partial y}{\partial r} \]

or in a convenient shorthand
\[ u_r = u_x x_r + u_y y_r = u_x \cos \theta + u_y \sin \theta \]
The other three are
\[ u_{\theta} = u_x x_{\theta} + u_y y_{\theta} = u_x (-r \sin \theta) + u_y (r \cos \theta)  \]
\[ v_r = v_x x_r + v_y y_r = v_x (\cos \theta) + v_y (\sin \theta)  \]
\[ v_{\theta} = v_x x_{\theta} + v_y y_{\theta} = v_x (-r \sin \theta) + v_y (r \cos \theta) \]

Our key insight is to use the relations given by the CRE in Cartesian coordinates
\[ u_x = v_y \]
\[ u_y = -v_x \]

Thus, starting with the first expression for partial derivatives above
\[ u_r = u_x \cos \theta + u_y \sin \theta \]

Use the CRE in $x,y$ to substitute in terms of $v$:
\[ u_r = v_y \cos \theta + (-v_x) \sin \theta \]
we can see that this is different from the fourth expression above
\[ v_{\theta} = v_x (-r \sin \theta) + v_y (r \cos \theta) \]
only by a factor of $r$:
\[ r u_r = v_{\theta} \]

which is what needed to prove.

The other one is:
\[ u_{\theta} = u_x (-r \sin \theta) + u_y (r \cos \theta)  \]
\[ = v_y (-r \sin \theta) - v_x (r \cos \theta)  \]
compare with
\[ v_r =  v_x (\cos \theta) + v_y (\sin \theta)  \]

We need a factor of $-r$:
\[ u_{\theta} = -r v_r \]

Like the original Cartesian version but with an extra factor of $r$ on the partials with respect to $r$.

\section*{polar derivative, more carefully}

To get the derivative, start with the version that we know for $x,y$ coordinates:
\[ f'(z) = u_x + i v_x \]

Our problem is to define $f'(z)$ in terms of $u_r$ and $v_r$.

Substitute for $u_x$ first.  Go back to the two equations involving $u_x$ above
\[ u_r = u_x \cos \theta + u_y \sin \theta \]
\[ u_{\theta} = u_x (-r \sin \theta) + u_y (r \cos \theta) \]

Multiply the first by $\cos \theta$
\[ u_r \cos \theta = u_x \cos^2 \theta + u_y \sin \theta \cos \theta \]
 and the second by $- \sin \theta/r$
\[ - \frac{u_{\theta}}{r} \sin \theta = u_x  \sin^2 \theta - u_y \sin \theta \cos \theta \]
add
\[ u_x = u_r \cos \theta - \frac{u_{\theta}}{r} \sin \theta \]
Substitute $u_{\theta}/r = - v_r$
\[ u_x = u_r \cos \theta + v_r \sin \theta \]

Now get the second and fourth equations, with $v_x$
\[ v_r = v_x \cos \theta + v_y \sin \theta \]
\[ v_{\theta} = v_x (-r \sin \theta) + v_y (r \cos \theta) \]

Multiply the first by $\cos \theta$
\[ v_r \cos \theta = v_x \cos^2 \theta + v_y \sin \theta \cos \theta \]
and the second by $-\sin \theta / r$:
\[ -\frac{v_{\theta}}{r} \sin \theta = v_x \sin^2 \theta - v_y \sin \theta \cos \theta \]
add
\[ v_x = v_r \cos \theta -\frac{v_{\theta}}{r} \sin \theta \]
Subtitute $v_{\theta}/r = u_r$
\[ v_x = v_r \cos \theta - u_r \sin \theta \]

Combine the two results:
\[ f'(z) = u_x + i v_x \]
\[ = u_r \cos \theta + v_r \sin \theta + i \ [ \ v_r \cos \theta - u_r \sin \theta \ ] \]
Group terms with $u_r$ and $v_r$ separately:
\[ = u_r (\cos \theta - i \sin \theta) + v_r (\sin \theta + i \cos \theta)  \]
Multiply the second term by $1 = - i \cdot i$
\[ = u_r (\cos \theta - i \sin \theta) + i v_r (- i \sin \theta + \cos \theta)  \]
\[ f'(z) = e^{-i \theta} \ [ \ u_r + i v_r \ ] \]

And since
\[ r u_r = v_{\theta} \]
\[ r v_r = -u_{\theta} \]
then
\[ f'(z) = \frac{1}{r} \ e^{-i \theta} \ [ \ v_{\theta} - i u_{\theta} \ ] \]
compare this with 
\[ f'(z) = v_y - iu_y \]
and notice that the factor in front is just $1/z$

\subsection*{example 1}

Let's see if we can do an example.  Suppose
\[ f(z) = \sqrt{z} \]
Written in terms of $r, \theta$ we have
\[ f(z) = \sqrt{r} e^{i \theta/2} \]
\[ = \sqrt{r} \cos \theta/2 + i \sqrt{r} \sin \theta/2 \]
Then
\[ u_r = \frac{\cos \theta/2}{2 \sqrt{r}} \]
\[ v_r = \frac{\sin \theta/2}{2 \sqrt{r}} \]
and
\[ \ [ \sqrt{z} \ ]' = \ [ \ \sqrt{r} e^{i \theta/2} \ ]' = e^{-i \theta} \ [ \ u_r + i v_r \ ] \]
\[ = \frac{1}{2 \sqrt{r}} \ [ \ e^{-i \theta} (\cos \theta/2 + i \sin \theta/2) \ ]    \]
\[ = \frac{1}{2 \sqrt{r}} e^{-i \theta/2} = \frac{1}{2 \sqrt{z}} \]

\subsection*{example 2}
Let's try
\[ f(z) = \frac{1}{z} = \frac{1}{r} e^{-i \theta} \]
\[ = \frac{1}{r} \cos - \theta + i \frac{1}{r} \sin - \theta \]
\[ = \frac{1}{r} \cos \theta - i \frac{1}{r} \sin \theta \]
So
\[ u_r = -\frac{1}{r^2} \cos \theta \]
\[ v_r = \frac{1}{r^2} \sin \theta \]
\[ f'(z) = e^{-i \theta}  (u_r + i v_r) \]
\[ = \frac{1}{r^2} ( e^{-i \theta}) (- \cos \theta + i \sin \theta) \]
\[ = \frac{1}{r^2} ( e^{-i \theta}) (- \cos - \theta - i \sin - \theta) \]
\[ = \frac{1}{r^2} ( e^{-i \theta}) (-1) (e^{-i \theta}) \]
\[ = -\frac{1}{r^2 e^{i 2\theta}} = -\frac{1}{z^2} \]

\end{document}