\documentclass[11pt, oneside]{article} 
\usepackage{geometry}
\geometry{letterpaper} 
\usepackage{graphicx}
	
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{parskip}
\usepackage{color}
\usepackage{hyperref}

\graphicspath{{/Users/telliott/Github/figures/}}

\title{Taylor series}
\date{}

\begin{document}
\maketitle
\Large

%[my-super-duper-separator]

\url{https://math.mit.edu/~jorloff/18.04/notes/topic7.pdf}

A standard geometric series with a finite number of terms is
\[ S_n = a(1 + r + r^2 + r^3 + \dots + r^n) \]
\[ S_n = a + ar + ar^2 + ar^3 + \dots + ar^n) \]
where $a$ is the \emph{first term} and $r$ is the \emph{common ratio}.

Because the series is \emph{finite}, we know the sum exists, which allows us to write $S_n$.  The standard trick is
\[ S_n \ r = ar + ar^2 + ar^3 + \dots + ar^n + ar^{n+1}) \]
\[ S_n (1 - r) = ar - ar^{n+1} \]
\[ S_n = \frac{a(1 - r^{n+1})}{1 - r} \]

A series is understood to have an infinite number of terms.
\[ S = a + ar + ar^2 + ar^3 + \dots = \lim_{n \rightarrow \infty} S_n \]
Provided that $|r| < 1$, the term $r^{n+1}/(1-r)$ goes to zero in the limit, and then
\[ S = \frac{a}{1-r} \]

\subsection*{expansions}
We will often see the expression 
\[ \frac{1}{w - z} \]

$w$ is a complex variable and $z$ is a complex number.

This looks a lot like the sum of a geometric series and can be converted to one in two different ways.
\[ \frac{1}{w - z} = \frac{1}{w} \cdot \frac{1}{1 - z/w} \]
The second term on the right-hand side is a geometric series in $z/w$:
\[ =  \frac{1}{w} (1 + \frac{z}{w} + (\frac{z}{w})^2 + (\frac{z}{w})^3 + \dots ) \]
which converges when $|z/w| < 1$, that is, when $|w| > |z|$.

The other way is
\[ \frac{1}{w - z} = -\frac{1}{z} \cdot \frac{1}{1 - w/z} \]
The second term on the right-hand side is a geometric series in $w/z$:
\[ =  -\frac{1}{z} (1 + \frac{w}{z} + (\frac{w}{z})^2 + (\frac{w}{z})^3 + \dots ) \]
which converges when $|w/z| < 1$, that is, when $|w| < |z|$.

\subsection*{power series in $z$}

Here is a power series in $z$, where $z_0$ is a fixed point:
\[ f(z) = \sum_{n=0}^{\infty} a_n(z - z_0)^n \]

Theorem (without proof):

There is a number $R \ge 0$ such that

$\circ$ \ if $R > 0$, then the series converges absolutely to an analytic function, for $|z - z_0| < R$.

$\circ$ \ The series diverges for $|z - z_0| > R$. $R$ is called the radius of convergence. The disk $|z - z_0| < R$ is called the disk of convergence.

$\circ$ \ The derivative is given by term-by-term differentiation.  The series for $f'$ has the same radius of convergence.

$\circ$ \ If $\gamma$ is a bounded curve inside the disk of convergence, the integral is given by term-by-term integration:
\[ \int_{\gamma} f(z) \ dz = \sum_{n=0}^{\infty} \ \int_{\gamma} a_n(z - z_0)^n \ dz \]

If $R = \infty$ the function $f(z)$ is called \emph{entire}.  Often, $R$ may be determined by the ratio test.

\subsection*{ratio test}
Consider the series $\sum_0^{\infty} c_n$.  If $L = \lim_{n \rightarrow \infty} |c_{n+1}/c_n|$ exists, then:

$\circ$ \ if $L < 1$ the series converges absolutely.

$\circ$ \ if $L > 1$ the series diverges.

$\circ$ \ if $L = 1$ we don't know.

\subsection*{examples}

Consider the geometric series
\[ 1 + z + z^2 + z^3 + \dots \]

The ratio of the absolute values of consecutive terms is $|z|$ and the limit is:
\[ L = \lim_{n \rightarrow \infty} \frac{|z+1|}{|z|} = |z| \]

Convergence follows when $L < 1$, that is when $|z| < 1$.

Consider
\[ f(z) = \sum_{n=0}^{\infty} \ \frac{z^n}{n!} \]

\[ L =  \lim_{n \rightarrow \infty} \frac{|z^{n+1}|}{(n+1) |z^n|} =  \lim_{n \rightarrow \infty} \frac{|z|}{n+1} \]

Convergence follows when $L < 1$, this happens for all values of $z$.

Of course, this is the exponential function $f(z) = e^z$, which is \emph{entire}.  It is analytic for the whole complex plane.

To summarize, 

\begin{quote}we've seen that a power series converges to an analytic function inside its disk of convergence. Taylor’s theorem completes the story by giving the converse: around each point of analyticity an analytic function equals a convergent power series.\end{quote}

\subsection*{power series}

Power series are at once a big complication and yet the source of the most important results in complex function theory.  There is just no avoiding them.

For a review of Taylor series for real variables, I've included a chapter from the calculus book (\hyperref[sec:taylor_real]{\textbf{here}}).

All analytic functions can be expanded as power series around a fixed point $z_0$.  Churchill says (sect. 44):

\begin{quote}Suppose that a function $f$ is analytic [i.e. has a derivative] throughout an open disk $|z - z_0| < R_0$, centered at $z_0$ and with radius $R_0$.  Then, at each point in that disk, $f(z)$ has the series representation\end{quote}

\[ f(z) = \sum_{n=0}^{\infty} \ a_n(z - z_0)^n \]

This is a Taylor series.  The cofactors are obtained from the corollary to Cauchy's integral formula

The general rule is:
\[ f^n(z) = \frac{n!}{2 \pi i} \int_C \frac{f(w)}{(w - z)^{n+1}} \ dw \]

We obtain:
\[ a_n = \frac{1}{2 \pi i} \int_{\gamma} \frac{f(z)}{(z - z_0)^{n+1}} \ dz = \frac{f^n (z_0)}{n!} \]

\subsection*{example}

The next bit is from HELM:

\url{https://learn.lboro.ac.uk/archive/olmp/olmp_resources/pages/wbooks_fulllist.html}

Taylor series have terms of the form
\[ a_n (x-x_0)^n \] 
where the series is summed over positive integers from $n = 0 \rightarrow \infty$ and the coefficients are
\[ a_n = \frac{f^{(n)}}{n!} \]
the nth derivative of $f$ divided by $n!$

As an example consider
\[ f(x) = \frac{1}{1 - x} \]
This has a singularity at $x = 1$.  We can get the Taylor series expanded around $0$ for this function (this special form is called the Maclaurin series).
\[ f(x) = 1 + x + x^2 + x^3 + \dots \]
We can show that this series is equal to what we started with
\[ \frac{1}{1 - x} = 1 + x + x^2 + x^3 + \dots \]
by multiplying the right-hand side by $(1-x)$.  Imagine two long rows of numbers, one the series itself, and the second containing all the terms of the series multiplied by $-x$.  It's clear that everything cancels except the term $1$.

Alternatively, we can take derivatives and construct the series formally:
\[ f(x) = \frac{1}{1 - x} = (1-x)^{-1} \]
\[ f'(x) = \frac{1}{(1 - x)^2} = (1-x)^{-2} \]
Notice that the minus sign from the exponent cancels the minus sign from the term $(1-x)$ obtained by the chain rule.  
\[ f''(x) = \frac{2}{(1 - x)^3} \]
\[ f'''(x) = \frac{3!}{(1 - x)^4} \]
and so on.

Now, evaluated at $x_0 = 0$, these derivatives are seen to collapse to just the factorial, so we construct the terms of the series as
\[ a_n = \frac{f^{(n)}}{n!} \]
\[ = n! \ \frac{1}{n!} \]
and the factorials also cancel.  This leaves the particularly simple form:
\[ \sum_{n=0}^{\infty} x^n = 1 + x + x^2 + x^3 + \dots \]

\subsection*{convergence}
For most series the big question is:  what is the radius of convergence?  The one we just wrote is clearly divergent for say, $x = 2$, and in fact for $|x| \ge 1$.

The series expansion for real functions is centered around a fixed point $x_0$ with terms like $(x - x_0)^n$, and the series has a finite sum, only converges for $x$ sufficiently close to $x_0$.

\[ |x - x_0| < r \]

Likewise, for complex functions, series expansions will usually only be valid for a circle (or disk, or region) of convergence in the Argand plane with 
\[ | z - z_0 | < R \]

Convergence can be decided by certain tests including the ratio test and the root test (but sometimes the result is not clear).

See the previous chapter.

As the source says:  

"One of the shortcomings of Taylor series is that the circle of convergence is often only a part of the region in which $f(z)$ is analytic.  The Laurent series is an attempt to represent $f(z)$ as a series at as many points as possible. We expand the series around a point of singularity up to, but not including, the singularity itself."

Laurent series involve an annulus, usually called $D$, which is a circle that has an empty small circle in its center, like a slice through a donut.

\subsection*{proof}

From \url{https://math.mit.edu/~jorloff/18.04/notes/topic7.pdf}

The statement of Taylor's theorem is that if $f(z)$ is analytic in a region $A$ and $z_0 \in A$ then
\[ f(z) = \sum_{n=0}^{\infty} a_n(z - z_0)^n \]

where the series converges on any disk $|z - z_0|$ contained in $A$.

We also have the coefficients as
\[ a_n = \frac{f^n(z_0)}{n!} = \frac{1}{2 \pi i} \int_{\gamma} \frac{f(z)}{(z - z_0)^{n+1}} \ dz \]

Proof.

We fix $0 < r_1 < r_2 < r$, and let $\gamma$ be the circle $|w - z_0| = r_2$ traversed counterclockwise.

The disk of convergence extends to the closest boundary of $A$ (at $r$), but $r_1$ and $r_2$ can be arbitrarily close to $r$.

As preparation, we note that for $w$ on $\gamma$ (at $r_2$) and $|z - z_0| < r_1$, we have that
\[ |z - z_0| < r_1 < r_2 = |w - z_0| \]

Therefore
\[ \frac{|z - z_0|}{|w - z_0|} < 1 \]

so we can write this
\[ \frac{1}{w - z} = \frac{1}{w - z_0} \cdot \frac{1}{1 - \frac{z-z_0}{w - z_0}} \]

Then the second factor can be expressed as a convergent geometric series giving:
\[ \frac{1}{w - z_0} \ \sum_{n=0}^{\infty} (\frac{z - z_0}{w - z_0})^n  = \sum_{n=0}^{\infty} \frac{(z - z_0)^n}{(w - z_0)^{n+1}} \]

Now use Cauchy's formula to write:
\[ f(z) = \frac{1}{2 \pi i} \ \int_{\gamma} \frac{f(w)}{w - z} \ dw \]
\[ = \frac{1}{2 \pi i} \ \int_{\gamma} \sum_{n=0}^{\infty} \frac{(z - z_0)^n}{(w - z_0)^{n+1}} f(w) \ dw \] 
We can switch the integral of sums to the sum of integrals
\[ = \sum_{n=0}^{\infty} \ [ \ \frac{1}{2 \pi i} \ \int_{\gamma} \frac{1}{(w - z_0)^{n+1}} f(w) \ dw \ ]  \ (z - z_0)^n \] 
and then use Cauchy's formula for derivatives (the corollary):
\[ = \sum_{n=0}^{\infty} \ \frac{f^n(z_0)}{n!} \ (z - z_0)^n \]

$\square$

Comparison of the last two expressions gives what we had for the formula for the coefficients where 
\[ a_n = \frac{f^n(z_0)}{n!} = \frac{1}{2 \pi i} \int_{\gamma} \frac{f(w)}{(w - z_0)^{n+1}} \ dw \]

\end{document}