\documentclass[11pt, oneside]{article} 
\usepackage{geometry}
\geometry{letterpaper} 
\usepackage{graphicx}
	
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{parskip}
\usepackage{color}
\usepackage{hyperref}

\graphicspath{{/Users/telliott/Github/figures/}}

\title{Taylor series}
\date{}

\begin{document}
\maketitle
\Large

%[my-super-duper-separator]

Power series are at once a big complication and yet the source of the most important results in complex function theory.  There is just no avoiding them.

For a review of Taylor series for real variables, I've included a chapter from the calculus book (\hyperref[sec:taylor_real]{\textbf{here}}).

All analytic functions can be expanded as power series around a fixed point $z_0$.  Churchill says (sect. 44):

\begin{quote}Suppose that a function $f$ is analytic [i.e. has a derivative] throughout an open disk $|z - z_0| < R_0$, centered at $z_0$ and with radius $R_0$.  Then, at each point in that disk, $f(z)$ has the series representation\end{quote}

\[ f(z) = \sum_{n=0}^{\infty} \ a_n(z - z_0)^n \]

This is a Taylor series.  The cofactors can be obtained from the corollary to Cauchy's integral formula.

The general rule is:
\[ f^n(z) = \frac{n!}{2 \pi i} \int_C \frac{f(w)}{(w - z)^{n+1}} \ dw \]

We obtain:
\[ a_n = \frac{1}{2 \pi i} \int_{\gamma} \frac{f(z)}{(z - z_0)^{n+1}} \ dz = \frac{f^n (z_0)}{n!} \]

\subsection*{example}

The next bit is from HELM:

\url{https://learn.lboro.ac.uk/archive/olmp/olmp_resources/pages/wbooks_fulllist.html}

Taylor series have terms of the form
\[ a_n (x-x_0)^n \] 
where the series is summed over positive integers from $n = 0 \rightarrow \infty$ and the coefficients are
\[ a_n = \frac{f^{(n)}}{n!} \]
the nth derivative of $f$ divided by $n!$

As an example consider
\[ f(x) = \frac{1}{1 - x} \]
This has a singularity at $x = 1$.  We can get the Taylor series expanded around $0$ for this function (this special form is called the Maclaurin series).
\[ f(x) = 1 + x + x^2 + x^3 + \dots \]
We can show that this series is equal to what we started with
\[ \frac{1}{1 - x} = 1 + x + x^2 + x^3 + \dots \]
by multiplying the right-hand side by $(1-x)$.  Imagine two long rows of numbers, one the series itself, and the second containing all the terms of the series multiplied by $-x$.  It's clear that everything cancels except the term $1$.

Alternatively, we can take derivatives and construct the series formally:
\[ f(x) = \frac{1}{1 - x} = (1-x)^{-1} \]
\[ f'(x) = \frac{1}{(1 - x)^2} = (1-x)^{-2} \]
Notice that the minus sign from the exponent cancels the minus sign from the term $(1-x)$ obtained by the chain rule.  
\[ f''(x) = \frac{2}{(1 - x)^3} \]
\[ f'''(x) = \frac{3!}{(1 - x)^4} \]
and so on.

Now, evaluated at $x_0 = 0$, these derivatives are seen to collapse to just the factorial, so we construct the terms of the series as
\[ a_n = \frac{f^{(n)}}{n!} \]
\[ = n! \ \frac{1}{n!} \]
and the factorials also cancel.  This leaves the particularly simple form:
\[ \sum_{n=0}^{\infty} x^n = 1 + x + x^2 + x^3 + \dots \]

\subsection*{convergence}
For most series the big question is:  what is the radius of convergence?  The one we just wrote is clearly divergent for say, $x = 2$, and in fact for $|x| \ge 1$.

The series expansion for real functions is centered around a fixed point $x_0$ with terms like $(x - x_0)^n$, and the series has a finite sum, only converges for $x$ sufficiently close to $x_0$.

\[ |x - x_0| < r \]

Likewise, for complex functions, series expansions will usually only be valid for a circle (or disk, or region) of convergence in the Argand plane with 
\[ | z - z_0 | < R \]

Convergence can be decided by certain tests including the ratio test and the root test (but sometimes the result is not clear).

See the previous chapter.

As the source says:  

"One of the shortcomings of Taylor series is that the circle of convergence is often only a part of the region in which $f(z)$ is analytic.  The Laurent series is an attempt to represent $f(z)$ as a series at as many points as possible. We expand the series around a point of singularity up to, but not including, the singularity itself."

Laurent series involve an annulus, usually called $D$, which is a circle that has an empty small circle in its center, like a slice through a donut.

\subsection*{proof}

From \url{https://math.mit.edu/~jorloff/18.04/notes/topic7.pdf}

The statement of Taylor's theorem is that if $f(z)$ is analytic in a region $A$ and $z_0 \in A$ then
\[ f(z) = \sum_{n=0}^{\infty} a_n(z - z_0)^n \]

where the series converges on any disk $|z - z_0|$ contained in $A$.

We also have the coefficients as
\[ a_n = \frac{f^n(z_0)}{n!} = \frac{1}{2 \pi i} \int_{\gamma} \frac{f(z)}{(z - z_0)^{n+1}} \ dz \]

Proof.

We fix $0 < r_1 < r_2 < r$, and let $\gamma$ be the circle $|w - z_0| = r_2$ traversed counterclockwise.

The disk of convergence extends to the closest boundary of $A$ (at $r$), but $r_1$ and $r_2$ can be arbitrarily close to $r$.

As preparation, we note that for $w$ on $\gamma$ (at $r_2$) and $|z - z_0| < r_1$, we have that
\[ |z - z_0| < r_1 < r_2 = |w - z_0| \]

Therefore
\[ \frac{|z - z_0|}{|w - z_0|} < 1 \]

so we can write this
\[ \frac{1}{w - z} = \frac{1}{w - z_0} \cdot \frac{1}{1 - \frac{z-z_0}{w - z_0}} \]

Then the second factor can be expressed as a convergent geometric series giving:
\[ \frac{1}{w - z_0} \ \sum_{n=0}^{\infty} (\frac{z - z_0}{w - z_0})^n  = \sum_{n=0}^{\infty} \frac{(z - z_0)^n}{(w - z_0)^{n+1}} \]

Now use Cauchy's formula to write:
\[ f(z) = \frac{1}{2 \pi i} \ \int_{\gamma} \frac{f(w)}{w - z} \ dw \]
\[ = \frac{1}{2 \pi i} \ \int_{\gamma} \sum_{n=0}^{\infty} \frac{(z - z_0)^n}{(w - z_0)^{n+1}} f(w) \ dw \] 
We can switch the integral of sums to the sum of integrals
\[ = \sum_{n=0}^{\infty} \ [ \ \frac{1}{2 \pi i} \ \int_{\gamma} \frac{1}{(w - z_0)^{n+1}} f(w) \ dw \ ]  \ (z - z_0)^n \] 
and then use Cauchy's formula for derivatives (the corollary):
\[ = \sum_{n=0}^{\infty} \ \frac{f^n(z_0)}{n!} \ (z - z_0)^n \]

$\square$

Comparison of the last two expressions gives what we had for the formula for the coefficients where 
\[ a_n = \frac{f^n(z_0)}{n!} = \frac{1}{2 \pi i} \int_{\gamma} \frac{f(w)}{(w - z_0)^{n+1}} \ dw \]

\subsection*{Constructing Taylor series}

We have
\[ f(z) = \sum_{n=0}^{\infty} \ a_n(z - z_0)^n \]
\[ a_n = \frac{1}{2 \pi i} \int_{\gamma} \frac{f(z)}{(z - z_0)^{n+1}} \ dz = \frac{f^n (z_0)}{n!} \]

For simplicity, just take $z_0 = 0$.
\[ f(z) = \sum_{n=0}^{\infty} \ a_n(z)^n \]
\[ a_n = \frac{1}{2 \pi i} \int_{\gamma} \frac{f(z)}{z^{n+1}} \ dz = \frac{f^n (0)}{n!} \]

To find the coefficients we need derivatives $f^n$.

The classic examples are, of course, the exponential and the sine and cosine.  The derivatives for complex functions are exactly analogous to those for real functions, so the Taylor formulas are too.

If $f(z) = e^z$ then all the derivatives $f^n$ are also $e^z$, and evaluated at $0$ give $1$.  So the series is
\[ e^z = \sum_{n=0}^{\infty} \ \frac{z^n}{n!} = 1 + z + \frac{z^2}{2!} + \frac{z^3}{3!}  \dots \]

If $f(z) = \sin z$ then the function and the even derivatives are zero at $z = 0$, while the odd ones are $\pm \cos z$, and evaluated at $0$ give $\pm 1$.  So the series is
\[ \sin z =  z - \frac{z^3}{3!} + \frac{z^5}{5!}  \dots \]

We can also use the definition of the complex sine to get this:
\[ \sin z = \frac{1}{2i} \ (e^{iz} - e^{-iz}) \]
Remember the factor of $1/2i$ and write
\[ = [1 + (iz) + \frac{(iz)}{2!}^2 + \frac{(iz)}{3!}^3 \dots \ ] \  - \ [ \ 1 + (-iz) + \frac{(-iz)}{2!}^2 + \frac{(-iz)}{3!}^3 \dots \ ]  \]
\[ = [1 + iz - \frac{z}{2!}^2 - i \frac{z^3}{3!} \dots \ ] \  - \ [ \ 1 - iz - \frac{z^2}{2!} + i \frac{z^3}{3!} \dots \ ]  \]
\[ = 2iz -  2i \frac{z^3}{3!} \dots  \]
We will get $2$ times $i$ times the odd powers of $z$ (over the factorial), and then an additional factor of $i^2$ causes a sign change for each term.  Multiply through by $1/2i$ to obtain
\[ 1 - \frac{z^3}{3!} + \frac{z^5}{5!} \dots \]

\subsection*{example (Orloff 7.12)}

Find the Taylor series for 
\[ f(z) = \frac{1}{1 - z} \]
around $z = 5$.  There is a bit of manipulation required to get the standard form:
\[ \frac{1}{1 - z} = \frac{1}{-4(1 + (z-5)/4)} \]

This is a geometric series with a leader factor of $-1/4$, and the power is of $-(z - 5)/4$.  That minus sign will give a minus sign to the odd powers in the expansion:
\[ -\frac{1}{4} \ [ \ 1 - (\frac{z-5}{4}) + (\frac{z-5}{4})^2 - (\frac{z-5}{4})^3 \dots \]

The series is constructed around $z = 5$, but the original function has a singularity at $z = 1$.  So the radius of convergence is the distance between these two.  Which matches
\[ |\frac{z-5}{4}| < 1 \]
\[ |z-5| < 4 \]

\subsection*{example (Orloff 7.13)}

Find the Taylor series for 
\[ f(z) = \log(1 + z) \]

The function $\log w$ is undefined for the real number line from $- \infty \rightarrow 0$.  Because of that one, the region where it's not defined extends up to $z = -1$.  We are asked to construct a Taylor series around $z=0$.  The radius of convergence will be $R = 1$.

The way to do this is to take the derivative of the function, construct its series, and then integrate term by term:
\[ f'(z) = \frac{1}{1 + z} = 1 - z + z^2 - z^3 \dots \]
The integral is
\[ f(z) = a_0 + z - \frac{z^2}{2} + \frac{z^3}{3} \dots \]

Find $a_0$ by evaluating at $z = 0$.  then
\[ f(z) = \log 1 = 0 = a_0 + 0 + 0 \dots \]
So $a_0$ equals zero.


\end{document}